---
title: "Qualitative Activity Recognition"
author: "Shevarev, M.V."
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: [Human Activity Recognition](http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har).  
*see the section on the Weight Lifting Exercise Dataset.

### Data

The training data for this project are available here:  
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:  
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source:   http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

```{r, echo=FALSE, results = "hide", warning=FALSE,  message=FALSE} 
# Load the necessary libraries
library(caret)
library(randomForest)
```

### Cleaning the data

```{r}
#download the data
url1 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
url2 <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(url1, "training.csv")
download.file(url2, "testing.csv")
trainig <- read.csv("training.csv")
testing <- read.csv("testing.csv")

#get information about the data
dim(trainig)
names(trainig)

#remove columns that have NA values more than 0.5 of total values
ncol <- colSums(is.na(trainig))<nrow(trainig)/2
trainig <- trainig[,ncol]
testing <- testing[,ncol]
emptycol <- colSums(trainig!="")>nrow(trainig)/2
trainig <- trainig[,emptycol]
testing <- testing[,emptycol]

#remove the non-informative features
trainig <- trainig[,-c(1:7)]
testing <- testing[,-c(1:7)]

#filter out highly correlated predictors
filter<-preProcess(trainig, method = "corr")
trainig <- predict(filter, trainig)
testing <- predict(filter, testing)

#the data set in this step
dim(trainig)
```

### Training and prediction

```{r}
#split the data into training and validation
set.seed(1337)
inTrain <- createDataPartition(trainig$classe, p = 0.6, list = FALSE)
Training <- trainig[inTrain,]
Testing <- trainig[-inTrain,]

#using randomforest algorithm and 3 folds
rf <- trainControl(method="cv", number=3, verboseIter=FALSE)
trainingModel <- train(classe ~ ., data=Training, method="rf",trControl=rf)
trainingModel$finalModel

#prediction
predict <- predict(trainingModel, Testing)
cm <-confusionMatrix(factor(Testing$classe), factor(predict))

#the validation accuracy
cm$overall['Accuracy']

#error
1 - cm$overall['Accuracy']
```

### Use the Prediction Model on the validation data
```{r}
predict(trainingModel, testing)
```